Leadership across cultures and contexts: an exploratory analysis

### 2020-05-03 leadership_across_cultures_contexts.Rmd
Tweaked some variable names. Fixed some figure/table captions.

---
title: "Leadership across cultures and contexts: an exploratory analysis"
author: "Zachary H. Garfield, Kristen L. Syme, and Edward H. Hagen"
date: "4/25/2019"
output: 
  bookdown::html_document2
bibliography: Leader2.bib
csl: evolution-and-human-behavior.csl
---

@@

There is now well established consensus within the behavioral sciences on the importance of data from the full range of human cultural diversity. Systemic biases in much of social science are well known, and might be due, in part, to the lack of data from populations other than American undergraduates [@henrich_weirdest_2010, *The weirdest people in the world?*]. Since the seminal WEIRD people paper, many social scientists have taken up the challenge to diversify their samples and have sought out "non-WEIRD" participants and databases. Dichotomizing populations and their psychologies and social dynamics as WEIRD or non-WEIRD, however, is a mistake.^[A Google Scholar search for "non-WEIRD" produced over 1000 results.] As anthropology has conclusively demonstrated, there is enormous diversity across societies that does not remotely resemble a dichotomy.

@@

There is potentially as much diversity within nonindustrial, non-WEIRD populations as there is within WEIRD ones, and there is also substantial overlap of the behaviors of people everywhere [@brown_human_1991]. @henrich_weirdest_2010 in their critique of the over-reliance on WEIRD samples emphasized a holistic approach to cultural diversity within and across "telescoping" levels of human societies. It is therefore important to assess cultural variation without assuming that everyone within a culture is essentially the same, and without putting undue emphasis on either the between-population differences or similarities.

@@

<!-- CLOSING -->
We first briefly review theoretical, ethnographic, and empirical perspectives on the qualities of leaders, the functions they serve, and the costs and benefits of leadership across human societies, including WEIRD and non-WEIRD populations. We then outline the aims and methods of the current cross-cultural study -- which focuses on uncovering patterns among non-WEIRD populations -- before presenting results and interpretations. Through novel exploratory analyses encompassing the full range of cultural diversity and diverse social contexts we aim to advance leadership studies across disciplines and provide researchers with a foundation for more detailed approaches to human leadership across cultures and contexts.

## The functions of leadership across cultures and contexts
<!-- INTRO -->
Despite the universality of the phenomenon of leadership, the function leaders serve across populations and social contexts is highly diverse and are expected to be closely tied to group structure and group goals [@van_vugt_cognitive_2007]. Among many of the ethnographically described small-scale societies, which generally lack multi-level, institutionalized social and political structures, leadership at the community level is typically informal and ephemeral and leaders emerge contextually when demands of coordination or conflict resolution require a managerial presence [@von_rueden_leadership_2015; @kantner_identifying_2010; @lewis_leaders_1974; @boehm_egalitarian_1993].

The theoretical and ethnographic literature suggests these demands most commonly arise in the contexts of collective actions and in conflict resolution [e.g., @glowacki_leadership_2015; @gavrilets_convergence_2016; @hooper_theory_2010; @von_rueden_solving_2015; @henrich_big_2015; @pietraszewski_evolution_2019]. Demands of mediation in within-group conflict resolution, although variable in scale, are probably a universal feature of human groups [@king_origins_2009; @van_vugt_leadership_2008]. In the absence of institutionalized organizational structure leaders often take on essential roles in within-group coordination and collective action, including facilitating conflict resolution within or between autonomous households or kin groups, managing individual investments to group-wide projects, and administering sanctions to individuals who fail to appropriately contribute to collective actions, comply with group policies, or seriously violate social norms. Demands of between-group interactions, including both conflict and cooperation are also very common across small-scale societies, including hunter-gatherers [@glowacki_evolutionary_2017; @hames_pacifying_2019; @richerson_cultural_2016; @henrich_evolution_2009; @apicella_social_2012]. Across relatively small-scale societies representing the group and managing both internal and external relationships and conflicts are expected to be the most salient functions of community leaders. Within communities, however, leaders may emerge whenever individuals congregate. The role of leaders in ritual contexts, such as in initiations [@singh_cultural_2017], and in providing social services, including facilitating marriages [@walker_evolutionary_2011] and organizing social activities [@wiessner_food_1996], are also expected to be common functions carried out by a leader.

@glowacki_leadership_2015 draw on empirical, ethnographic, and theoretical models of leader emergence and suggest in small-scale societies institutionalized leadership roles are likely to initially emerge to resolve conflicts between families. @roscoe_social_2009 draws on the ethnography of highland New Guinea in describing the modular structure of human societies and suggests the functions of community leaders in small-scale societies are structured by the joint demands of signaling military formidability to rival groups and facilitating within-group conflict resolution and cooperation. For "big-men" in these societies, in addition to military prowess, "talents for oratory, mediation, and organization were instrumental in reducing their internal conflict and enhancing the ability of their members to coordinate their actions as a group" [@roscoe_social_2009, p. 104].

<!-- GETS INTO QUALITIES BELOW -->
<!-- Given demands of community organization and conflict resolution, leadership tends to be context-specific and leaders in such contexts often draw on persuasive capacities and leverage community-based respect, specialized skills or knowledge, and social support to achieve or maintain community influence [@kantner_identifying_2010; @lewis_leaders_1974; @boehm_egalitarian_1993].  -->

@@

Our second goal was to extract higher-order dimensions of leader qualities and functions from our researcher coded variables. We used several methods to explore lower dimensional representations of these data that might reveal informative "features." These included logistic PCA, non-negative matrix factorization (NMF), and hierarchical cluster analysis. Logistic PCA generalizes Pearson's original formulation of PCA to find a projection of the natural parameters of the saturated model that minimizes the Bernoulli deviance [@logisticPCA].

@@

With an interpretable method of dimensionality reduction, we aimed to predict variation in our "feature" dimensions from population-level variation, such as subsistence type, region, and group structure.

@@

Our third goal was to determine which leader and follower benefits and costs were most commonly mentioned in the ethnographic record, computing proportions and 95% CIs as above and to estimate how costs and benefits varied by group structure types.

@@

The analyses computing levels of evidence for each variable used the entire set of text records. Analyses investigating lower dimensional representations used a slightly smaller set that omitted records with no evidence for any variable (i.e., all zeros). All analyses were conducted with `r R.version.string`.

@@

The geographic distribution of the cultures in our sample is displayed in Figure \@ref(fig:geographic), with cultures labeled by subsistence strategy and represented by a symbol proportional to the number of documents that provided ethnographic texts for that culture (see also Supplementary Figure \@ref(fig:subsisttree)). Our search yielded `r nrow(all_data)` text records from `r length(unique(leadershipdata::documents$d_ID))` unique documents describing `r length(unique(all_data$d_culture))` cultures. The number of text records per culture ranged from `r min(leader_cult$number_leader_records, na.rm=T)` to `r max(leader_cult$number_leader_records, na.rm=T)`, with a median of `r median(leader_cult$number_leader_records, na.rm=T)`, a mean of `r signif(mean(leader_cult$number_leader_records, na.rm=T), 3)`, and a standard deviation of `r signif(sd(leader_cult$number_leader_records, na.rm=T), 3)`. The text records were generally short, ranging from `r textstats$min` to `r textstats$max` words, with a median word count of `r textstats$median`, a mean of `r textstats$mean`, and standard deviation of `r textstats$sd`. Of the `r nrow(all_data)` text records, `r final_record_count` had evidence for at least one variable in the study.

@@

For each of the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` coded variables (see Tables \@ref(tab:qualfuncvars) and \@ref(tab:costbenefitvars)), we computed the proportion of text records (95% CI) that provided support using logistic mixed effects models that controlled for one possible source of bias -- publication date -- as a fixed effect, and included authors nested within cultures as random intercepts. We then computed the proportion of cultures with at least one text record providing support for that variable (calculating 95% CI using a cluster bootstrap). See Figure \@ref(fig:varsupport). For a table of all coefficients and 95% CIs, see Table S\@ref(tab:glmmcoefs).

@@

To investigate variation in functions, qualities, costs, and benefits, we fit a multiple logistic mixed effects regression model of each variable as a function of subsistence strategy, continental region, leader sex, and group structure, with random intercepts for author nested within culture. We then compared the AIC value of each model to a model with only the fixed and random intercepts, retaining `r nrow(multi_aic)` models that showed at least a two-point reduction in AIC value. We then created heatmaps that displayed the relative estimated marginal means for each level of these categorical explanatory variables for each outcome variable. See Figures S\@ref(fig:heatmapsubsis), S\@ref(fig:heatmapregion), S\@ref(fig:heatmapsex), and S\@ref(fig:heatmapgroups), and Tables S\@ref(tab:variablevars) and S\@ref(tab:universalvars).

@@

```{r leadertotalbenefits, fig.width=10, fig.cap="Estimated marginal means of A. leader total benefits, and B. total costs, as a function of group structure, based on a logistic mixed effects regression model with random intercepts for ethnographer nested within culture. Values are on the response scale."}
# p1 <- 
#   hagenutils::ggemmeans(emmeans(m_ldrtotben, type = 'response', specs = 'group.structure2')) +
#   scale_x_continuous(limits = c(0, 0.1)) +
#   labs(title = "Total leader benefits", x = "", y = "")
# 
# p2 <- hagenutils::ggemmeans(emmeans(m_ldrtotcost, type = 'response', specs = 'group.structure2')) +
#   scale_x_continuous(limits = c(0, 0.1)) +
#   labs(title = "Total leader costs", x = "\nProbability", y = "")
# 
# p1 + p2 + plot_layout(ncol = 1) + plot_annotation(tag_levels = 'A')

@@

```{r leadertotalcosts}
theplots <- all_emms(m_ldrtotcost, c('subsistence', 'group.structure2'), 0.06, 'Predictors of leader costs')

@@

```{r groupstatus, fig.width=10, fig.cap="Estimated marginal means of a mixed effects logistic regression model of evidence for high status (0/1) as a function of group type, subsistence strategy, continental region, and leader sex, with random effects for ethnographer nested within culture. Values are on the response scale (probability)."}
# ((plot_emm_status_group2 / plot_emm_status_sex2) | (plot_emm_status_subsistence2 / plot_emm_status_region2)) + plot_annotation(title = "Predictors of high status", tag_levels = 'A')

@@

## Subsistence strategy and group context effects on human leadership

@@

Across cultures and contexts, when ethnographic texts described female leaders exclusively, they were relatively more likely to be described as older, of high status, married to a high-quality spouse in a polygynous marriage, and the recipients of various mating benefits. This suite of characteristics is consistent with discussions of high status first-wives who gain and maintain increased social influence across the lifespan by leveraging their extended kin and social networks [@goodale_tiwi_1971; @brown_her_1985], and supports the view of positive assortative mating of high status leaders @garfield_investigating_2019 documented among elected leaders in a small-scale society with relative gender-egalitarianism. Evidence for male leaders (relative to evidence for female leaders or evidence that was not sex-specific) was indeed linked to group-level and political functions and male leaders were relativey more likely to provide social services, control political structure, and were at greater risk of losing their high social standing. Female leaders, relative to males, were also more likely to lack coercive authority, which is consistent with much of leadership studies suggesting dominant and aggressive women are less likely to be preferred as leaders, unlike dominent men who are sometimes prefered as leaders [@garfield_investigating_2019; @williams_subtle_2016]. See Figure S\@ref(fig:heatmapsex).

Despite sex differences, there is also likely to be substantial overlap in many of the qualities and functions of female and male leaders across human societies. Leader sex was not a predictor of the majority of the measures investigated here (adjusting for other population-level measures; see Table S\@ref(tab:universalvars)). When sex was a predictor (Table S\@ref(tab:variablevars)), descriptions including both female and male leaders provided relatively more supporting evidence for many key measures of leader qualities and functions, including group organization, resource distribution, ritual functions, and being brave and experienced and accomplished )Figure S\@ref(fig:heatmapsex).

@@

```{r heatmapregion, fig.cap="Heatmap of relative estimated marginal means of the levels of evidence for each outcome variable by each subsistence strategy. Estimated marginal means were standardized by rows. The colors therefore represent the relative amount of evidence within the same outcome variable."}

@@

```{r heatmapsex, fig.cap="Heatmap of relative estimated marginal means of the levels of evidence for each outcome variable by each subsistence strategy. Estimated marginal means were standardized by rows. The colors therefore represent the relative amount of evidence within the same outcome variable."}

@@

```{r heatmapgroups, fig.cap="Heatmap of relative estimated marginal means of the levels of evidence for each outcome variable by each subsistence strategy. Estimated marginal means were standardized by rows. The colors therefore represent the relative amount of evidence within the same outcome variable."}

@@

```{r variablevars}

multi_aic %>% 
  select(
    var,
    Variable, 
    Sum, 
    uniAIC, 
    multiAIC, 
    AIC_diff,
    pvalue_subsis,
    pvalue_region,
    pvalue_female,
    pvalue_groups
    ) %>% 
  mutate(
    var = map_chr(var, ~ str_split(., '_')[[1]][[1]])
  ) %>% 
  arrange(Variable) %>% 
  kable(
  caption = 'Retained logistic mixed effects models of each variable as functions of subsistence strategy, continental region, leader sex, and group structures. Models were retained due to improvement in AIC values relative to an intercept-only model. Sum: the number of rows with evidence for the variable. uniAIC: AIC of intercept-only model. multiAIC: AIC of multivariate model. AIC_diff: difference in AIC values. P-values are the unadjusted p-values of each explanatory variable in the model.',
  digits = 2
  )
```

```{r universalvars}
multi_allvars %>%
  dplyr::filter(AIC_diff >= -2) %>% 
  select(
    var,
    Variable, 
    Sum, 
    uniAIC, 
    multiAIC, 
    AIC_diff,
    pvalue_subsis,
    pvalue_region,
    pvalue_female,
    pvalue_groups
    ) %>% 
  mutate(
    var = map_chr(var, ~ str_split(., '_')[[1]][[1]])
  ) %>% 
  arrange(Variable) %>% 
  kable(
  caption = 'Logistic mixed effects models that did not show improvement in AIC values relative to an intercept-only model. These study variables either did not vary much by subsistence strategy, continental region, leader sex, or group structures, or had little evidence. Sum: the number of rows with evidence for the variable. uniAIC: AIC of intercept-only model. multiAIC: AIC of multivariate model. AIC_diff: difference in AIC values. P-values are the unadjusted p-values of each explanatory variable in the model.',
  digits = 2
  )

@@

We constructed `r length(features)` "feature" variables from the higher-level clusters (Figure \@ref(fig:cluster)): `r paste(features, collapse = ', ')`. The Mediate, Prosociality, and Organize features were modestly positively correlated with each other, but not with Prestige or Social/material success. Big man was correlated with Cultural conformity. See Figure \@ref(fig:featurecor).

>>>>

---
title: "Leadership across cultures and contexts: an exploratory analysis"
author: "Zachary H. Garfield, Kristen L. Syme, and Edward H. Hagen"
date: "4/25/2019"
output: 
  bookdown::html_document2:
    number_sections: false
bibliography: Leader2.bib
csl: evolution-and-human-behavior.csl
---

@@

There is now well established consensus within the behavioral sciences on the importance of data from the full range of human cultural diversity. Systemic biases in much of social science are well known, and might be due, in part, to the lack of data from populations other than American undergraduates [@henrich_weirdest_2010, *The weirdest people in the world?*].

Since the seminal WEIRD people paper, many social scientists have taken up the challenge to diversify their samples and have sought out "non-WEIRD" participants and databases. Dichotomizing populations and their psychologies and social dynamics as WEIRD or non-WEIRD, however, is a mistake.^[A Google Scholar search for "non-WEIRD" produced over 1000 results.] As anthropology has conclusively demonstrated, there is enormous diversity across societies that does not remotely resemble a dichotomy.

@@

There is potentially as much diversity within nonindustrial, non-WEIRD cultures as there is within WEIRD ones, and there is also substantial overlap of the behaviors of people everywhere [@brown_human_1991]. @henrich_weirdest_2010 in their critique of the over-reliance on WEIRD samples emphasized a holistic approach to cultural diversity within and across "telescoping" levels of human societies. It is therefore important to assess cultural variation without assuming that everyone within a culture is essentially the same, and without putting undue emphasis on either the between-population differences or similarities.

@@

<!-- CLOSING -->
We first briefly review theoretical, ethnographic, and empirical perspectives on the qualities of leaders, the functions they serve, and the costs and benefits of leadership across human societies, including WEIRD and non-WEIRD populations. We then outline the aims and methods of the current cross-cultural study before presenting results and interpretations. Through novel exploratory analyses encompassing the full range of cultural diversity and diverse social contexts we aim to advance leadership studies across disciplines and provide researchers with a foundation for more detailed approaches to human leadership across cultures and contexts.

<!-- Reference a historical example of a short term leader implementing change with lasting effects, good and/or bad. -->

## The functions of leadership across cultures and contexts
<!-- INTRO -->
Despite the universality of the phenomenon of leadership, the function leaders serve across populations and social contexts is highly diverse and are expected to be closely tied to group structure and group goals [@van_vugt_cognitive_2007]. Among many of the ethnographically described small-scale societies, which generally lack multi-level, institutionalized social and political structures, leadership at the community level is typically informal and ephemeral and leaders emerge contextually when demands of coordination or conflict resolution require a managerial presence [@von_rueden_leadership_2015].

<!-- FUNCTIONS in SSS = CORDINATION AND CONFCLIT RESOLUTION -->
In the absence of institutionalized organizational structure leaders often take on essential roles in group coordination and collective action, including facilitating conflict resolution within or between autonomous households or kin groups, managing individual investments to group-wide projects, and administering sanctions to individuals who do not appropriately contribute to collective actions, comply with group policies, or seriously violate social norms. @glowacki_leadership_2015 draw on empirical, ethnographic, and theoretical models of leader emergence and suggest in small-scale societies institutionalized leadership roles are likely to initially emerge to resolve conflicts between families. @roscoe_social_2009 draws on the ethnography of highland New Guinea in describing the modular structure of human societies and suggests the functions of community leaders in small-scale societies are structured by the joint demands of signaling military formidability to rival groups and facilitating within-group conflict resolution and cooperation. For "big-men" in these societies, in addition to military prowess, "talents for oratory, mediation, and organization were instrumental in reducing their internal conflict and enhancing the ability of their members to coordinate their actions as a group" [@roscoe_social_2009, p. 104].

<!-- GETS INTO QUALITIES BELOW -->
Given demands of community organization and conflict resolution, leadership tends to be context-specific and leaders in such contexts often draw on persuasive capacities and leverage community-based respect, specialized skills or knowledge, and social support to achieve or maintain community influence [@kantner_identifying_2010; @lewis_leaders_1974; @boehm_egalitarian_1993].

<!-- REPETITIVE COMPARED TO ABOVE BUT INCLUDES ADDITIONAL DETAILS about FUNCTIONS AND  MORE CITATIONS, THEORY -->
The theoretical and ethnographic literature suggests these demands most commonly arise in the contexts of collective actions and in conflict resolution [e.g., @glowacki_leadership_2015; @gavrilets_convergence_2016; @hooper_theory_2010; @von_rueden_solving_2015; @henrich_big_2015; @pietraszewski_evolution_2019]. Demands of mediation in within-group conflict resolution, although variable in scale, are probably a universal feature of human groups [@king_origins_2009; @van_vugt_leadership_2008]. Demands of between-group interactions, including both conflict and cooperation are also very common across small-scale societies, including hunter-gatherers [@glowacki_evolutionary_2017; @hames_pacifying_2019; @richerson_cultural_2016; @henrich_evolution_2009; @apicella_social_2012]. Across relatively small-scale societies representing the group and managing both internal and external relationships and conflicts are expected to be the most salient functions of community leaders. Within communities, however, leaders may emerge whenever individuals congregate. The role of leaders in ritual contexts, such as in initiations [@singh_cultural_2017], and in providing social services, including facilitating marriages [@walker_evolutionary_2011] and organizing social activities [@wiessner_food_1996], are also expected to be common functions carried out by a leader.

@@

Our second goal was to determine which leader and follower benefits and costs were most commonly mentioned in the ethnographic record, computing proportions and 95% CIs as above and to estimate how costs and benefits varied by group structure types.

Our third goal was to assess the extent to which each of the leadership trait and cost/benefit variables varied by four theoretically important factors for which we had data for all rows: subsistence strategy, continental region, group context, and leader sex.

Our fourth goal was to extract higher-order dimensions of leader qualities and functions from our researcher coded variables. We used several methods to explore lower dimensional representations of these data that might reveal informative "features." These included logistic PCA, non-negative matrix factorization (NMF), and hierarchical cluster analysis. Logistic PCA generalizes Pearson's original formulation of PCA to find a projection of the natural parameters of the saturated model that minimizes the Bernoulli deviance [@logisticPCA].

@@

Our fifth goal was to assess variation in our "feature" dimensions by subsistence strategy, continental region, group context, and leader sex.

@@

@@

The analyses computing levels of evidence for each variable used the entire set of text records. Analyses investigating lower dimensional representations used a smaller set that omitted rows with no evidence for any variable (i.e., all zeros), and variables with very little evidence (almost all zeros). All analyses were conducted with `r R.version.string`.

@@

The geographic distribution of the cultures in our sample is displayed in Figure \@ref(fig:geographic), with cultures labeled by subsistence strategy and represented by a symbol proportional to the number of documents that provided ethnographic texts for that culture (see also Supplementary Figure S\@ref(fig:subsisttree)). Our search yielded `r nrow(all_data)` text records from `r length(unique(leadershipdata::documents$d_ID))` unique documents describing `r length(unique(all_data$d_culture))` cultures. The number of text records per culture ranged from `r min(leader_cult$number_leader_records, na.rm=T)` to `r max(leader_cult$number_leader_records, na.rm=T)`, with a median of `r median(leader_cult$number_leader_records, na.rm=T)`, a mean of `r signif(mean(leader_cult$number_leader_records, na.rm=T), 3)`, and a standard deviation of `r signif(sd(leader_cult$number_leader_records, na.rm=T), 3)`. The text records were generally short, ranging from `r textstats$min` to `r textstats$max` words, with a median word count of `r textstats$median`, a mean of `r textstats$mean`, and standard deviation of `r textstats$sd`. Of the `r nrow(all_data)` text records, `r final_record_count` had evidence for at least one variable in the study.

@@

For each of the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` coded variables (see Tables \@ref(tab:qualfuncvars) and \@ref(tab:costbenefitvars)), we computed the proportion of text records (95% CI) that provided support using intercept-only logistic mixed effects models that included authors nested within cultures as random intercepts. We then computed the proportion of cultures with at least one text record providing support for that variable (calculating 95% CI using a cluster bootstrap). See Figure \@ref(fig:varsupport). For a table of all coefficients and 95% CIs, see Table S\@ref(tab:glmmcoefs).

@@

To investigate variation in functions, qualities, costs, and benefits, we fit a multiple logistic mixed effects regression model of each variable as a function of subsistence strategy, continental region, leader sex, and group structure, with random intercepts for author nested within culture (a total of `r nrow(multi_allvars)` models). We then compared the AIC value of each model to a model with only the fixed and random intercepts, retaining `r nrow(multi_aic)` models that showed at least a two-point reduction in AIC value. To compare the relative effects of each level of each categorical predictor variable on each outcome variable, we created heatmaps that displayed the relative estimated marginal means for each level of these categorical explanatory variables for each outcome variable. See Figures S\@ref(fig:heatmapsubsis), S\@ref(fig:heatmapregion), S\@ref(fig:heatmapsex), and S\@ref(fig:heatmapgroups), and Tables S\@ref(tab:variablevars) and S\@ref(tab:universalvars).

@@

```{r leadertotalbenefits, fig.width=10, fig.cap="Estimated marginal means of a logistic mixed effects regression model of leader total benefits as a function of A. subsistence strategy, B. group structure, C. continental region, and D. leader sex. Model included random intercepts for ethnographer nested within culture. Values are on the response scale."}
# p1 <- 
#   hagenutils::ggemmeans(emmeans(m_ldrtotben, type = 'response', specs = 'group.structure2')) +
#   scale_x_continuous(limits = c(0, 0.1)) +
#   labs(title = "Total leader benefits", x = "", y = "")
# 
# p2 <- hagenutils::ggemmeans(emmeans(m_ldrtotcost, type = 'response', specs = 'group.structure2')) +
#   scale_x_continuous(limits = c(0, 0.1)) +
#   labs(title = "Total leader costs", x = "\nProbability", y = "")
# 
# p1 + p2 + plot_layout(ncol = 1) + plot_annotation(tag_levels = 'A')

@@

```{r leadertotalcosts, fig.cap="Estimated marginal means of a logistic mixed effects regression model of leader total costs as a function of A. subsistence strategy, and B. group structure. Model included random intercepts for ethnographer nested within culture. Values are on the response scale."}
theplots <- all_emms(m_ldrtotcost, c('subsistence', 'group.structure2'), 0.06, 'Predictors of leader costs')

@@

```{r groupstatus, fig.width=10, fig.cap="Estimated marginal means of a mixed effects logistic regression model of evidence for high status (0/1) as a function of A. subsistence strategy, B. group structure, C. continental region, and D. leader sex. Model included random intercepts for ethnographer nested within culture. Values are on the response scale (probability)."}
# ((plot_emm_status_group2 / plot_emm_status_sex2) | (plot_emm_status_subsistence2 / plot_emm_status_region2)) + plot_annotation(title = "Predictors of high status", tag_levels = 'A')

@@

## Culture, context, and human leadership

@@

Across cultures and contexts, when ethnographic texts described female leaders exclusively, they were relatively more likely to be described as older, of high status, married to a high-quality spouse in a polygynous marriage, and the recipients of various mating benefits. This suite of characteristics is consistent with discussions of high status first-wives who gain and maintain increased social influence across the lifespan by leveraging their extended kin and social networks [@goodale_tiwi_1971; @brown_her_1985], and supports the view of positive assortative mating of high status leaders @garfield_investigating_2019 documented among elected leaders in a small-scale society with relative gender-egalitarianism. Evidence for male leaders (relative to evidence for female leaders or evidence that was not sex-specific) was indeed linked to group-level and political functions and male leaders were relativey more likely to provide social services, control political structure, and were at greater risk of losing their high social standing. See Figure S\@ref(fig:heatmapsex).

Female leaders were also more likely to lack coercive authority, which is consistent with much of leadership studies suggesting dominant and aggressive women are less likely to be preferred as leaders, unlike dominent men who are sometimes prefered as leaders [@garfield_investigating_2019; @williams_subtle_2016].

Despite sex differences, there is also likely to be substantial overlap in many of the qualities and functions of female and male leaders across human societies. Leader sex was not a predictor of the majority of the measures investigated here (adjusting for other population-level measures; see Table S\@ref(tab:universalvars)). When sex was a predictor (Table S\@ref(tab:variablevars)), descriptions including both female and male leaders provided relatively more supporting evidence for many key measures of leader qualities and functions, including group organization, resource distribution, ritual functions, and being brave and experienced and accomplished.

@@

```{r heatmapregion, fig.cap="Heatmap of relative estimated marginal means of the levels of evidence for each outcome variable by each continental region. Estimated marginal means were standardized by rows. The colors therefore represent the relative amount of evidence within the same outcome variable."}

@@

```{r heatmapsex, fig.cap="Heatmap of relative estimated marginal means of the levels of evidence for each outcome variable by leader sex. Estimated marginal means were standardized by rows. The colors therefore represent the relative amount of evidence within the same outcome variable."}

@@

```{r heatmapgroups, fig.cap="Heatmap of relative estimated marginal means of the levels of evidence for each outcome variable by each group context. Estimated marginal means were standardized by rows. The colors therefore represent the relative amount of evidence within the same outcome variable."}

@@

```{r multiout}
multi_out <-
  multi_allvars %>% 
  select(
    var,
    Variable, 
    Evidence, 
    interceptAIC, 
    multiAIC, 
    AIC_diff,
    pvalue_subsis,
    pvalue_region,
    pvalue_sex,
    pvalue_groups
    ) %>% 
  mutate(
    var = map_chr(var, ~ str_split(., '_')[[1]][[1]])
  ) %>% 
  arrange(Variable)
```

```{r variablevars}
multi_out %>% 
  dplyr::filter(AIC_diff < -2) %>% 
  kable(
  caption = 'Retained logistic mixed effects models of each variable as functions of subsistence strategy, continental region, leader sex, and group structures. Models were retained if their AIC values decreased by two or more relative to an intercept-only model. Evidence: the number of rows with evidence for the variable. interceptAIC: AIC of intercept-only model. multiAIC: AIC of multivariate model. AIC_diff: difference in AIC values. P-values are the unadjusted p-values of each predictor variable in the model.',
  digits = 2
  )
```

```{r universalvars}
multi_out %>%
  dplyr::filter(AIC_diff >= -2) %>% 
  kable(
  caption = 'Logistic mixed effects models that did not show improvement in AIC values relative to an intercept-only model. These study variables either did not vary much by subsistence strategy, continental region, leader sex, or group structures, or had little evidence. Evidence: the number of rows with evidence for the variable. interceptAIC: AIC of intercept-only model. multiAIC: AIC of multivariate model. AIC_diff: difference in AIC values. P-values are the unadjusted p-values of each predictor variable in the model.',
  digits = 2
  )

@@

We constructed `r length(features)` "feature" variables from the higher-level clusters (Figure \@ref(fig:cluster)): `r paste(features, collapse = ', ')`. The Mediate, Prosociality, and Organize features were modestly positively correlated with each other, but not with Prestige or Social/material success. Prosocial competencies was correlated with Cultural conformity. See Figure \@ref(fig:featurecor).
